---
title: "Final Project"
description: |
  Permentation Test Across March Madness Dataset
author: "Chloe Denhart"
date: December 9, 2024
format: 
  revealjs:
    fontSize: 15px 
execute: 
  warning: false
  message: false
---

## Overview

1.  Project 1: Netflix Analysis
2.  Project 2: March Madness Permutation Test
3.  Project 3: WAI Database

## Netflix Analysis

-   Categorized Netflix titles by genre.
-   Explored relationships between genres, release years, and description lengths.

## Figure 1 - Movies per Genre

```{r}
#| echo: false
netflix_titles <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-04-20/netflix_titles.csv')
```

```{r}
#| echo: false
library(tidyverse)
library(ggplot2)
library(tidytuesdayR)
library(stringr)
```

```{r}
#| echo: false
netflix_titles2 <- netflix_titles |>
  mutate(listed_in = str_replace_all(listed_in, "[A-Z]", tolower))
```

```{r}
#| echo: false
genre_titles <- netflix_titles2 |>
  filter(str_detect(listed_in, "comedy|comedies|drama|horror|documentary|action|crime|reality")) |>
  mutate(genre = case_when(
    str_detect(listed_in, "comedy|comedies") ~ "Comedy",
    str_detect(listed_in, "drama") ~ "Drama",
    str_detect(listed_in, "horror") ~ "Horror",
    str_detect(listed_in, "documentary") ~ "Documentary",
    str_detect(listed_in, "action") ~ "Action",
    str_detect(listed_in, "crime") ~ "Crime",
    str_detect(listed_in, "reality") ~ "Reality"
  )) |>
  select(genre, type)
```

```{r}
#| echo: false
year_counts <- netflix_titles |>
  group_by(release_year, type) |>
  summarise(count = n()) |>
  arrange(desc(count))
```

```{r}
#| echo: false

genre_plot <- ggplot(genre_titles, aes(x = genre)) +
    geom_bar() +
    labs(
      x = "Genre",
      y = "Count",
      title = "# of Movies per Genre"
    )
print(genre_plot)

```

## Figure 2 - Number of Titles for TV Shows and Movies

```{r}
#| echo: false
year_plot <- 
  ggplot(year_counts, aes(x = release_year , y = count, color = type)) +
  geom_point() +
  labs(title = "Number of Titles", 
       x = "Year", 
       y = "Count"
       )
print(year_plot)
```

## March Madness

-   Conducted a permutation test to analyze upset probabilities in 2023.
-   Explored the upset rate for lower-seeded teams in the round of 64.
-   Null hypothesis (H0) : the probability that a lower seeded team wins being less than or equal to 90% (Upset win rate ≤ 90%).
-   Alternative hypothesis (Ha): the probability that a lower seeded team wins being greater than 90% (Upset win rate \> 90%)

## Outcome of Upsets

-   This tibble shows how many upsets occurred throughout the round of 64

```{r}
#| echo: false
march_madness <- read.csv("538 Ratings.csv")

```

```{r}
#| echo: false
library(dplyr)
library(ggplot2)
library(tidyverse)
library(purrr)
```

```{r}
#| echo: false
total_games <- march_madness |>
  filter(YEAR == 2023, ROUND == 64) |>
  nrow()
```

```{r}
#| echo: false
upsets <- march_madness |>
  filter(YEAR == 2023, ROUND < 64, SEED > 8) |>
  nrow()
```

```{r}
#| echo: false
march_madness |>
  filter(YEAR == 2023) |>
  filter(ROUND <= 64) |>
  mutate(RESULT = ifelse(ROUND < 64, "win", "loss")) |>
  mutate(seed_group = ifelse(SEED <=8, "top", "bottom")) |>
  group_by(seed_group, RESULT) |>
  summarize(n())
# upset rate = 0.21875, which is 7/32
```

Upset rate (7/32):

```{r}
#| echo: false
# upset rate drawn from the total number of games in round of 64, and the amount of games won by a team when their seed >= 8. See table above.
upset_rate <- upsets/total_games
upset_rate
  
```

```{r}
#| echo: false
random_choice <- function(num_teams) {
  shuffled_seed <- sample(c("lower", "higher"), num_teams, replace = TRUE, prob = c(0.9, 0.1))
  upset_ratio <- mean(shuffled_seed == "higher")
  return(upset_ratio)
}
```

```{r mapping function}
#| echo: false
set.seed(47)
num_exper <- 5000
null_distribution <- map_dbl(1:num_exper, ~ random_choice(total_games))
# issue with rendering is that total games is not defined
```

P-Value:

```{r}
#| echo: false
p_value <- sum(null_distribution >= upset_rate) / num_exper
p_value
```

## Null Distribution

-   Histogram of null distribution with upset rate highlighted.

```{r}
#| echo: false
ggplot(data.frame(null_distribution), aes(x = null_distribution)) + 
  geom_histogram(bins = 30, color = "black", fill = "skyblue") +
  geom_vline(xintercept = upset_rate, color = "red") + # red line should be at upset rate not p-value
  labs(x = "Proportion of teams that upset a higher seed", 
       y = "Count",
       title = "Sampling distribution under the null hypothesis")
```

## Key Insights

-   Upset rate: 21.9%.
-   P-value: 0.035 — rejected null hypothesis. Lower-seeded teams can achieve significant upsets.

## WAI Database Analysis

-   Analyzed ear frequency and mean absorbency data from 12 studies.
-   Compared sex-based differences in absorbency for Aithal (2013).

## Figure 1 - Mean Absorbtion From 12 Studies

```{r}
#| echo: false
library(RMariaDB)
library(tidyverse)
library(ggplot2)
```

```{r}
#| echo: false
library(RMariaDB)
library(tidyverse)

library(RMariaDB)
con_wai <- dbConnect(
  MariaDB(), host = "scidb.smith.edu",
  user = "waiuser", password = "smith_waiDB", 
  dbname = "wai"
)
Measurements <- tbl(con_wai, "Measurements")
PI_Info <- tbl(con_wai, "PI_Info")
Subjects <- tbl(con_wai, "Subjects")

# collect(Measurements)

```

```{sql}
#| connection: con_wai
#| output.var: ears

#| echo: false
SELECT 
  m.Identifier,
  CONCAT(pi.AuthorsShortList,pi.Year,m.Instrument, m.Frequency) AS Label,
  m.Frequency,
  AVG(m.Absorbance) AS Mean_Absorbance,
  COUNT(DISTINCT m.SubjectNumber, m.Ear) AS Unique_Ears
FROM Measurements AS m
JOIN PI_Info AS pi
  ON m.Identifier = pi.Identifier
WHERE m.Identifier IN ('Abur_2014', 'Werner_2010', 'Voss_2020', 'Feeny_2017', 'Groon_2015', 'Lewis_2015', 'Liu_2008', 'Rosowski_2012', 'Shahnaz_2006', 'Shaver_2013', 'Sun_2016', 'Voss_1994', 'Voss_2010')
GROUP BY m.Identifier, m.Frequency, pi.AuthorsShortList, pi.Year, m.Instrument
ORDER BY pi.Year, m.Frequency;


```

```{r}
#| echo: false
ears |> 
  ggplot(aes(x = Frequency, y = Mean_Absorbance, color = Identifier)) +
  geom_line() +
  labs(
    title = "Mean Absorbance From Each Publication in WAI Database",
    x = "Frequency (Hz)",
    y = "Mean Absorbance",
    color = "Study"
  ) +
  scale_x_continuous(limits = c(200, 8000)) +
  scale_y_continuous(limits = c(0, 1)) +
  theme_minimal()


```

## Figure 2 - Analysis by Sex (using the *Aithal (2013)* study)

```{sql}
#| connection: con_wai
#| output.var: sex_analysis
#| echo: false
SELECT 
  m.Identifier,
  m.Frequency,
  s.Sex,
  AVG(m.Absorbance) AS Mean_Absorbance,
  COUNT(DISTINCT m.SubjectNumber, m.Ear) AS Unique_Ears
FROM Measurements AS m
JOIN Subjects AS s
  ON m.SubjectNumber = s.SubjectNumber
WHERE m.Identifier = 'Aithal_2013'
GROUP BY m.Frequency, s.Sex
ORDER BY m.Frequency, s.Sex;

```

```{r}
#| echo: false
sex_analysis |>
  ggplot(aes(x = Frequency, y = Mean_Absorbance, color = Sex)) +
  geom_line() +
  labs(
    title = "Frequency vs. Mean Absorbance by Sex for Aithal_2013",
    x = "Frequency (Hz)",
    y = "Mean Absorbance",
    color = "Sex"
  ) +
  theme_minimal() +
  scale_x_continuous(limits = c(200, 8000)) +
  scale_y_continuous(limits = c(0, 1)) +
  theme_minimal()


```

## Questions?
