---
title: "Mini-Project 2"
description: |
  Analysis of Netflix Movies and TV Shows
author: Chloe Denhart
date: October 2, 2024
format: html
execute: 
  warning: false
  message: false
---

In this project, I analyzed the Netflix Movies and TV Shows dataset to explore trends and insights into the platform's content. This dataset, sourced from Kaggle and made available through TidyTuesday, provides information on titles, genres, release years, ratings, and descriptions.

```{r}
netflix_titles <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-04-20/netflix_titles.csv')
```

```{r}
library(tidyverse)
library(ggplot2)
library(tidytuesdayR)
library(stringr)
```

```{r}
netflix_titles2 <- netflix_titles |>
  mutate(listed_in = str_replace_all(listed_in, "[A-Z]", tolower))
```

```{r}
genre_titles <- netflix_titles2 |>
  filter(str_detect(listed_in, "comedy|comedies|drama|horror|documentary|action|crime|reality")) |>
  mutate(genre = case_when(
    str_detect(listed_in, "comedy|comedies") ~ "Comedy",
    str_detect(listed_in, "drama") ~ "Drama",
    str_detect(listed_in, "horror") ~ "Horror",
    str_detect(listed_in, "documentary") ~ "Documentary",
    str_detect(listed_in, "action") ~ "Action",
    str_detect(listed_in, "crime") ~ "Crime",
    str_detect(listed_in, "reality") ~ "Reality"
  )) |>
  select(genre, type)

```

```{r}
genre_plot <-ggplot(genre_titles, aes(x = genre)) +
    geom_bar() +
    labs(
      x = "Genre",
      y = "Count",
      title = "# of Movies per Genre"
    )
print(genre_plot)
```

```{r}
library(plotly)

# Creating the Plotly interactive plot for genre distribution
genre_plotly <- genre_titles |>
  count(genre) |>
  plot_ly(x = ~genre, y = ~n, type = 'bar', 
          text = ~paste('Count: ', n), 
          hoverinfo = 'text') |>
  layout(title = "Number of Movies per Genre",
         xaxis = list(title = "Genre"),
         yaxis = list(title = "Count"))

# Display the interactive plot
genre_plotly

```

```{r}
year_counts <- netflix_titles |>
  group_by(release_year, type) |>
  summarise(count = n()) |>
  arrange(desc(count))

```

```{r}
year_plot <- 
  ggplot(year_counts, aes(x = release_year , y = count, color = type)) +
  geom_point() +
  labs(title = "Number of Titles by Genre", 
       x = "Genre", 
       y = "Count"
       )
print(year_plot)
```

```{r}
description_length <- netflix_titles |>
  mutate(description_length = str_length(description)) |>
  select(title, description, description_length) |>
  arrange(desc(description_length))
  
```

```{r}
mean_description_length <- description_length |>
  summarize(mean_length = mean(description_length, na.rm = TRUE))

mean_description_length
```

```{r}
rating <- netflix_titles |>
  filter(str_detect(rating, "^(R|PG-13|PG|G)$")) |>
  select(title, rating)
rating
```

**Analysis on Tables/Plots:**

My first table, I wanted to change the genre descriptions to all lowercase. I did this so that for my next table I could condense all of the genres into six different categories, allowing me to plot a graph showing the amount of movies/television shows in each genre.

My first graph represents that amount of movies/television shows in each genre. From this graph, we can see the most common genres are comedy and drama. The least common genres are crime and reality.

I upgraded my first graph by producing a interactive plot with plotly.

From my third graph we can see the relationship between year, and movie/television show release date. We can infer from this graph that that a lot more movies and television shows were put on Netflix after the mid 2000s.

I wanted to play around with the str_length function. So I took the length of all of the descriptions of each title. The mean of the description lengths was 143.1004 string characters.

Lastly, I created a new data frame that takes the title of the show and detects the R, PG-13, PG and G ratings

**Reference:**

Bansal, Shivam. *Netflix Movies and TV Shows Dataset*. Kaggle.\
Dataset available at [Netflix Shows Dataset](https://www.kaggle.com/datasets/shivamb/netflix-shows).

*Netflix Shows* from Kaggle, contains information about various movies and TV shows available on Netflix. The dataset includes attributes such as the title, genre, language, release year, and rating of each title. It provides a snapshot of Netflix's catalog and allows for various analyses, such as trends in content release, genre popularity, or the evolution of Netflix's original content over time.
