---
title: "Mini-Project 4"
description: |
  Analysis of Ear Frequencies and Mean Absorbancy
author: Chloe Denhart
date: November 21, 2024
format: html
execute: 
  warning: false
  message: false
---

In this project, I will reproduce the graph created in Figure 1 of the *Voss (2020)* study using data from the Wideband Acoustic Immittance (WAI) Database. The graph will plot frequency against mean absorption for the 12 studies referenced in Figure 1, showing the differences across studies. I will also explore a second study, *Aithal (2013)*, focusing on differences in mean absorption by sex. Both analyses will involve querying the WAI Database with SQL to extract the necessary data and using R to create plots.

```{r}
library(RMariaDB)
library(tidyverse)
library(ggplot2)
```

```{r}
library(RMariaDB)
library(tidyverse)

library(RMariaDB)
con_wai <- dbConnect(
  MariaDB(), host = "scidb.smith.edu",
  user = "waiuser", password = "smith_waiDB", 
  dbname = "wai"
)
Measurements <- tbl(con_wai, "Measurements")
PI_Info <- tbl(con_wai, "PI_Info")
Subjects <- tbl(con_wai, "Subjects")

# collect(Measurements)

```

![]()

![]()

```{sql}
#| connection: con_wai

SHOW TABLES;

```

```{sql}
#| connection: con_wai

DESCRIBE Measurements;

```

```{sql}
#| connection: con_wai

SELECT *
FROM Measurements
LIMIT 0, 5;

```

```{sql}
#| connection: con_wai

SELECT DISTINCT Identifier
FROM Measurements;

```

```{sql}
#| connection: con_wai
#| output.var: ears
SELECT 
  m.Identifier,
  CONCAT(pi.AuthorsShortList,pi.Year,m.Instrument, m.Frequency) AS Label,
  m.Frequency,
  AVG(m.Absorbance) AS Mean_Absorbance,
  COUNT(DISTINCT m.SubjectNumber, m.Ear) AS Unique_Ears
FROM Measurements AS m
JOIN PI_Info AS pi
  ON m.Identifier = pi.Identifier
WHERE m.Identifier IN ('Abur_2014', 'Werner_2010', 'Voss_2020', 'Feeny_2017', 'Groon_2015', 'Lewis_2015', 'Liu_2008', 'Rosowski_2012', 'Shahnaz_2006', 'Shaver_2013', 'Sun_2016', 'Voss_1994', 'Voss_2010')
GROUP BY m.Identifier, m.Frequency, pi.AuthorsShortList, pi.Year, m.Instrument
ORDER BY pi.Year, m.Frequency;


```

```{r}
ears |> 
  ggplot(aes(x = Frequency, y = Mean_Absorbance, color = Identifier)) +
  geom_line() +
  labs(
    title = "Mean Absorbance From Each Publication in WAI Database",
    x = "Frequency (Hz)",
    y = "Mean Absorbance",
    color = "Study"
  ) +
  scale_x_continuous(limits = c(200, 8000)) +
  scale_y_continuous(limits = c(0, 1)) +
  theme_minimal()


```

This graphs analyzes 12 different studies from the WAI database. Each line represents data from a specific study, identified by the author, year, instrument, and the number of unique ears measured. The x-axis shows frequency in Hertz (Hz), ranging from 200 Hz to 8000 Hz, while the y-axis displays mean absorption values ranging from 0 to 1. This plot highlights variations in mean absorption across different studies and instruments, demonstrating how measurement techniques and populations may influence auditory outcomes. Each study has differing ranges in mean absorbency and frequency.

Now we will reproduce a similar graph, finding a study in the same database where subjects of different sex. In doing this I will be using the *Aithal (2013)* study.

```{sql}
#| connection: con_wai
#| output.var: sex_analysis
SELECT 
  m.Identifier,
  m.Frequency,
  s.Sex,
  AVG(m.Absorbance) AS Mean_Absorbance,
  COUNT(DISTINCT m.SubjectNumber, m.Ear) AS Unique_Ears
FROM Measurements AS m
JOIN Subjects AS s
  ON m.SubjectNumber = s.SubjectNumber
WHERE m.Identifier = 'Aithal_2013'
GROUP BY m.Frequency, s.Sex
ORDER BY m.Frequency, s.Sex;

```

```{sql}
#| connection: con_wai
#| output.var: sex

SELECT Sex, Identifier
FROM Subjects
```

```{r}
sex_analysis |>
  ggplot(aes(x = Frequency, y = Mean_Absorbance, color = Sex)) +
  geom_line() +
  labs(
    title = "Frequency vs. Mean Absorbance by Sex for Aithal_2013",
    x = "Frequency (Hz)",
    y = "Mean Absorbance",
    color = "Sex"
  ) +
  theme_minimal() +
  scale_x_continuous(limits = c(200, 8000)) +
  scale_y_continuous(limits = c(0, 1)) +
  theme_minimal()


```

This graphs analyzes just the *Aithal (2013)* study from the WAI database. Each line represents data from a specific sex group. There seems to be some mistake in the data where "Female" was spelled "Femake". The x-axis shows frequency in Hertz (Hz), ranging from 200 Hz to 8000 Hz, while the y-axis displays mean absorption values ranging from 0 to 1. This plot highlights that there are not distince variations in mean absorption and frequency across different sex's, demonstrating that sex does not influence auditory outcomes.
