---
title: "Text Analysis"
description: |
  Analysis of Netflix Movies and TV Shows
author: Chloe Denhart
date: October 2, 2024
format: html
execute: 
  warning: false
  message: false
---

In this project, I analyzed the Netflix Movies and TV Shows dataset to explore trends and insights into the platform's content. This dataset, sourced from Kaggle and made available through TidyTuesday, provides information on titles, genres, release years, ratings, and descriptions.

```{r}
netflix_titles <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-04-20/netflix_titles.csv')
```

```{r}
library(tidyverse)
library(ggplot2)
library(tidytuesdayR)
library(stringr)
```

```{r}
# changed all captitalized letters to lower in order to analyze
netflix_titles2 <- netflix_titles |>
  mutate(listed_in = str_replace_all(listed_in, "[A-Z]", tolower))
```

```{r}
# Chose 6 main genres to examine
genre_titles <- netflix_titles2 |>
  filter(str_detect(listed_in, "comedy|comedies|drama|horror|documentary|action|crime|reality")) |>
  mutate(genre = case_when(
    str_detect(listed_in, "comedy|comedies") ~ "Comedy",
    str_detect(listed_in, "drama") ~ "Drama",
    str_detect(listed_in, "horror") ~ "Horror",
    str_detect(listed_in, "documentary") ~ "Documentary",
    str_detect(listed_in, "action") ~ "Action",
    str_detect(listed_in, "crime") ~ "Crime",
    str_detect(listed_in, "reality") ~ "Reality"
  )) |>
  select(genre, type)

```

```{r}
genre_plot <-ggplot(genre_titles, aes(x = genre)) +
    geom_bar() +
    labs(
      x = "Genre",
      y = "Count",
      title = "# of Movies per Genre"
    )
print(genre_plot)
```

```{r}
library(plotly)

# Creating the Plotly interactive plot for genre distribution
genre_plotly <- genre_titles |>
  count(genre) |>
  plot_ly(x = ~genre, y = ~n, type = 'bar', 
          text = ~paste('Count: ', n), 
          hoverinfo = 'text') |>
  layout(title = "Number of Movies per Genre",
         xaxis = list(title = "Genre"),
         yaxis = list(title = "Count"))

# Display the interactive plot
genre_plotly

```

```{r}
# create data frame that counts amount of titles in within each year
year_counts <- netflix_titles |>
  group_by(release_year, type) |>
  summarise(count = n()) |>
  arrange(desc(count))

```

```{r}
year_plot <- 
  ggplot(year_counts, aes(x = release_year , y = count, color = type)) +
  geom_point() +
  labs(title = "Number of Titles by Genre", 
       x = "Genre", 
       y = "Count"
       )
print(year_plot)
```

```{r}
# calculates length of descriptions
description_length <- netflix_titles |>
  mutate(description_length = str_length(description)) |>
  select(title, description, description_length) |>
  arrange(desc(description_length))
  
```

```{r}
# calculates average length of descriptions
mean_description_length <- description_length |>
  summarize(mean_length = mean(description_length, na.rm = TRUE))

mean_description_length
```

```{r}
# tibble with just rating and movie title
rating <- netflix_titles |>
  filter(str_detect(rating, "^(R|PG-13|PG|G)$")) |>
  select(title, rating)
rating
```

**Analysis on Tables/Plots:**

My first table, I wanted to change the genre descriptions to all lowercase. I did this so that for my next table I could condense all of the genres into six different categories, allowing me to plot a graph showing the amount of movies/television shows in each genre.

My first graph represents that amount of movies/television shows in each genre. From this graph, we can see the most common genres are comedy and drama. The least common genres are crime and reality.

I upgraded my first graph by producing a interactive plot with plotly.

From my third graph we can see the relationship between year, and movie/television show release date. We can infer from this graph that that a lot more movies and television shows were put on Netflix after the mid 2000s.

I wanted to play around with the str_length function. So I took the length of all of the descriptions of each title. The mean of the description lengths was 143.1004 string characters.

Lastly, I created a new data frame that takes the title of the show and detects the R, PG-13, PG and G ratings

**Data Source**:\
Netflix Movies and TV Shows Dataset\

**Author**: Shivam Bansal

This data was sourced from Kaggle and contains information about various movies and TV shows available on Netflix. Specifically, it includes:

-   Attributes such as title, genre, language, release year, and rating for each movie or TV show.

-   The dataset offers insights into Netflix’s catalog, making it useful for analyses such as trends in content release, genre popularity, and the evolution of Netflix’s original content over time.

-   The data was last updated on Kaggle by Shivam Bansal.
